---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
if(!file.exists("activity.csv")){
    unzip("activity.zip")
}
activity <- read.csv("activity.csv")
activity$date <- as.Date(activity$date)
summary(activity)
```



## What is mean total number of steps taken per day?
1. Calculating the total number of steps taken per day 
```{r}
steps_by_day <- aggregate(steps~date , data = activity ,FUN = sum , na.rm = TRUE)
```
2. Make a histogram of the total number of steps taken each day
```{r}
hist(steps_by_day$steps , xlab="Number of steps" , ylab = "frequency" , main = "Total number of steps taken each day")
```
3. Mean and Median of the total number of steps taken per day
```{r}
meanbyday <- format(mean(steps_by_day$steps),digits = 1)
medianbyday <- format(median(steps_by_day$steps) , digits = 1)
```
**Mean** : `r meanbyday`
**Median** : `r medianbyday`

## What is the average daily activity pattern?
1. Make a time series plot (i.e. \color{red}{\verb|type = "l"|}type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
daily_activity <- aggregate(steps~interval,data = activity,FUN = mean , na.rm = TRUE)
plot(daily_activity$interval,daily_activity$steps,type = "l" , xlab = "5 Min. Intervals" , ylab = "Average steps taken" , main = "Average daily activity pattern",col = "steelblue")

```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps
```{r}
max_steps <- max(daily_activity$steps)
max_stepsinterval <- daily_activity$interval[which(daily_activity$steps==max_steps)]
```
**The interval Which contains maximum number of steps is : `r max_stepsinterval` **


## Imputing missing values
1. Calculating the total no. of missing values
```{r}
sum(is.na(activity$steps))
```
2. making the copy of data and replacing the NA's with mean
```{r}
copydata <- activity
nas <- is.na(copydata$steps)
avg_interval <- tapply(copydata$steps, copydata$interval, mean, na.rm=TRUE, simplify=TRUE)
copydata$steps[nas] <- avg_interval[as.character(copydata$interval[nas])]
```
3. 
```{r}
steps_per_day <- aggregate(steps~date , data = copydata ,FUN = sum , na.rm = TRUE)
hist(steps_per_day$steps , xlab="Number of steps" , ylab = "frequency" , main = "Total number of steps taken each day")
```
4. mean and median total number of steps taken per day
```{r}
mean <- format(mean(steps_per_day$steps),digits = 1)
median <- format(median(steps_per_day$steps) , digits = 1)
```
**Mean** : `r mean`
**Median** : `r median`

## Are there differences in activity patterns between weekdays and weekends?
1. Creating new factor variable
```{r}
library(dplyr)
copydata <- mutate(copydata, weektype = ifelse(weekdays(copydata$date) == "Saturday" | weekdays(copydata$date) == "Sunday", "weekend", "weekday"))
copydata$weektype <- as.factor(copydata$weektype)
head(copydata)
```
2. Plotting the graphs
```{r}
library(ggplot2)
interval_full <- copydata %>%
  group_by(interval, weektype) %>%
  summarise(steps = mean(steps))
s <- ggplot(interval_full, aes(x=interval, y=steps, color = weektype)) +
  geom_line() +
  facet_wrap(~weektype, ncol = 1, nrow=2)
print(s)
```
